Shallow EDSLs and Object-Oriented Programming

### 2016-06-14 paper/paper.tex
Format Scala code

\lstset{ %
language=Java,                % choose the language of the code
columns=fullflexible,
keepspaces=true,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbers=none,
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
aboveskip=3pt,
belowskip=-1pt
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}
\begin{document}

>>>>

\lstdefinelanguage{JavaScala}{
  morekeywords={public,int,interface,implements,default,
    abstract,case,void,catch,class,def,static,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,var,val,while,yield,with},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}

\lstset{ %
language=JavaScala,                % choose the language of the code
columns=fullflexible,
keepspaces=true,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbers=none,
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
aboveskip=3pt,
belowskip=-1pt
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}
\begin{document}

### 2016-06-14 paper/sections/ShallowOO.tex
Format Scala code

The rewrite goes on by using a record with a sole field to represent the result:
\lstinputlisting[language=haskell,linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
We no longer need a separate \texttt{width3} function as the field name is the extractor.
The implementation is still shallow since essentially it is nothing different
from the last one.
However, the definition already approaches an OO program such that we can easily
translate it into an OO language like Java:
\lstinputlisting[linerange=7-26]{../src/paper/sec2/Circuit.java}%APPLY:CIRCUIT_OO
where the record type maps to the abstract class \texttt{Circuit} and field
declaration becomes method declaration.
Each case in the semantic function corresponds to an class and its parameters become fields of the class.
And all the classes extends \texttt{Circuit} by giving \texttt{width} a definition.

@@

Implementing the desugaring method in Java could be:
\lstinputlisting[linerange=31-61]{../src/paper/sec2/Circuit.java}%APPLY:DESUGAR_OO
The newly introduced method \texttt{desugar} in \texttt{Circuit} returns a \texttt{Circuit} reflecting that it is a transformation. The implementation of \texttt{desugar} for each case is
straightforward: \emph{fan n} just returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
their results into an instance of \texttt{Beside}; \emph{identity n} mimics the
formula.

>>>>

The rewrite goes on by using a record with a sole field to represent the result:
\lstinputlisting[language=haskell,linerange=33-38]{./code/shallowCircuit.hs}%APPLY:CIRCUIT3
We no longer need a separate \texttt{width3} function as the field name is the extractor.
The implementation is still shallow since essentially it is nothing different
from the last one.
However, the definition already approaches an OO program such that we can easily
translate it into an OO language like Java:
\lstinputlisting[linerange=2-13]{../src/scala/base/Circuit.scala}%APPLY:CIRCUIT_SCALA
where the record type maps to the abstract class \texttt{Circuit} and field
declaration becomes method declaration.
Each case in the semantic function corresponds to an class and its parameters become fields of the class.
And all the classes extends \texttt{Circuit} by giving \texttt{width} a definition.

@@

Implementing the desugaring method in Java could be:
\lstinputlisting[linerange=3-20]{../src/scala/desugar/Circuit.scala}%APPLY:DESUGAR_SCALA
The newly introduced method \texttt{desugar} in \texttt{Circuit} returns a \texttt{Circuit} reflecting that it is a transformation. The implementation of \texttt{desugar} for each case is
straightforward: \emph{fan n} just returns itself; $beside\ c_1\ c_2$ recursively desugars $c_1$ and $c_2$ then wraps back
their results into an instance of \texttt{Beside}; \emph{identity n} mimics the
formula.
